---
title: Code Book - FORECASTING DEFINED CONTRIBUTION AND DEFINED BENEFIT RETIREMENT
  ACCOUNTS
author: "James Bishop, Hyungue Lim"
date: "March 16, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Things to note before proceeding
- Original data manipulations can be found in the excel file called "Documentation"

- Annual returns and volatility are calculated from S&P 500 values from January 1, 1950 - January 1, 2018

- Returns are yearly. Volatility of each year is calculated by taking standard deviation of monthly returns and multiplying by (12)^0.5

##Data Description
- contribution_4p: Includes contribution dollar amounts at age 25, 35, and 45. Initial salary of $50,000 at each age. Contribution percentage fixed at 4%. 
- contribution_8p: Includes contribution dollar amounts at age 25, 35, and 45. Initial salary of $50,000 at each age. Contribution percentage fixed at 8%. 
- contribution_graded: Includes contribution dollar amounts at age 25, 35, and 45. Initial salary of $50,000 at each age. Contribution percentage graded. 4% at age 25, 1% increase every 10 years there after.
- Service_Start_25 column: Contributions of initial service start age at 25
- Service_Start_35 column: Contributions of initial service start age at 35
- Service_Start_45 column: Contributions of initial service start age at 45
- In all functions, a denotes years left until retirement. b denotes service start age, age 25 being 0.
- For b, for instance, if service start age is 35, b=10
- For 4% contribution rate, use "contribution_4p" data frame
- For 8% contribution rate, use "contribution_8p" data frame
- For graded contribution rate, use "contribution_graded" data frame

#Code used for Fixed Interest Rate with 4% contribution and service start age at 25
```{r, contribution_4p}
Fixed <- function(a,b) {
  
    Future_Value <- rep(x=0, a)
    for (i in 1:a) {
      Future_Value[i] <- contribution_4p$Service_Start_25[i+b]*(1.0734)^(a-i+1)
    }
    final <- sum(Future_Value)
    
    print(mean(final))
}
Fixed(40,0) #Result for lumpsum at age 65
```

#Code used for Bootstrapping with 4% contribution and service start age at 25
```{r}
Bootstrapp <- function(a,b) {
  final <- rep(x=0, 1000)
  for (j in 1:1000) {
    
    Bootstrap_Return <- 0
    Bootstrap_Return <- sample(SnP500$Return, size = a, replace = TRUE)
    Bootstrap_Return <- as.data.frame(Bootstrap_Return)
    
    Future_Value <- rep(x=0, a)
  for (i in 1:a) {
    Future_Value[i] <- contribution_4p$Service_Start_25[i+b]*prod(Bootstrap_Return$Bootstrap_Return[i:a])
    
  }
  final[j] <- sum(Future_Value)
  }
  print(mean(final))
  print(sd(final))
}

Bootstrap(40,0) #Result for lumpsum at age 65
```

#Code used for Binomial Bootstrapping with 4% contribution and service start age at 25
```{r}
Binomial_Bootstrap <- function(a,b){
  final <- rep(x=0, 1000)
  for (k in 1:1000) {
    Expected_Value <- 0
    
    for (j in 1:a) { 
      Random_Return <- sample(1:68, 1, replace = TRUE)
      up <- 0; dn <- 0; prob <- 0
      up <- 1+((SnP500$Return[Random_Return]) + SnP500$Volatility[Random_Return])
      dn <- 1+((SnP500$Return[Random_Return]) - SnP500$Volatility[Random_Return])
      prob <- ((1+(SnP500$Return[Random_Return]) - dn) / (up - dn))
      
      #calculating tree. yes is initially 0, so adding contribution from the beginning is ok.
      Expected_Value = (Expected_Value + contribution_4p$Service_Start_25[j+b]) * up * prob + (Expected_Value+ contribution_4p$Service_Start_25[j+b]) * dn * (1 - prob)
      
      
    }
    final[k] <- Expected_Value
  }
  print(mean(final))
  print(sd(final))
}

Binomial_Bootstrap(40,0) #Result for lumpsum at age 65
```